---
title: "Homework 03: Use dplyr/ggplot2 to manipulate and explore data"
output: 
  github_document: default
  html_document: default
---
<style type="text/css">
.twoC {width: 100%}
.clearer {clear: both}
.twoC .table {max-width: 50%; float: right}
.twoC img {max-width: 50%; float: left}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```
```{r}
library(gapminder)
library(tidyverse)
library(gganimate)
```

#### Task #1 Get the maximum and minimum of GDP per capita for all continents.

```{r}
gdpPercap_continent <- gapminder %>% 
  group_by(continent) %>% 
  summarise(minGDP = min(gdpPercap), maxGDP = max(gdpPercap)) 

knitr::kable(gdpPercap_continent, col.names = c("Continnet", "min gdpPercap", "max gdpPercap"))
```

Then, plot the figures. 

##### The minimum gdpPercap for each continent: 
```{r}
gdpPercap_continent %>% 
  ggplot(aes(continent, minGDP, fill=continent)) + 
  geom_bar(stat="identity") +
  ggtitle("min gdpPercap for each continent")
```

##### The maximum gdpPercap for each continent:
```{r}
gdpPercap_continent %>% 
  ggplot(aes(continent, maxGDP, fill=continent)) + 
  geom_bar(stat="identity") +
  ggtitle("max gdpPercap for each continent")
```

#### Task #2 Look at the spread of GDP per capita within the continents.
Generate the spread of GDP per capita for each continent by calculating the minimum, mean, median, and maximum of gdpPercap

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarise(min = min(gdpPercap), mean = mean(gdpPercap), 
            median = median(gdpPercap), max = max(gdpPercap)) %>% 
  knitr::kable()
```

```{r}
gapminder %>% 
ggplot(aes(continent, log(gdpPercap), fill=continent)) + 
  geom_boxplot()
```

#### Task #3 Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.
I calcualte the weighted mean of life expectancy by pop for each continent and plot the change over time as followng:

```{r}
gapminder %>% 
  select(continent, year, lifeExp, pop) %>% 
  group_by(continent, year) %>% 
  summarise(mean = weighted.mean(lifeExp, pop)) 
```

```{r}
gapminder %>%
  group_by(continent, year) %>% 
  summarise(mean = weighted.mean(lifeExp, pop)) %>% 
  ggplot(aes(year, mean)) +
  geom_smooth(method="lm", se=TRUE) + # linear regression 
  geom_point(aes(colour = continent)) +
  ggtitle("Weighted Mean life expectancy changes over time for each continent")
```

#### Task #4 How is life expectancy changing over time on different continents?

I calcualte the mean of life expectancy for each continent and plot the change over time as followng:
```{r}
gapminder %>% 
  select(continent, year, lifeExp, pop) %>% 
  group_by(continent, year) %>% 
  summarise(mean = mean(lifeExp)) 
```

```{r}
gapminder %>%
  group_by(continent, year) %>% 
  summarise(mean = mean(lifeExp)) %>% 
  ggplot(aes(year, mean)) +
  geom_smooth(method="lm", se=TRUE) + # linear regression 
  geom_point(aes(colour = continent)) +
  ggtitle("Mean life expectancy changes over time for each continent")
```

To get better visualization of how the life expectancy changes over time on different continents, I used the `gganimate` library to generate dynamic figures (gif). 
```{r}
theme_set(theme_bw())  # pre-set the bw theme.
gapminder %>% 
  ggplot(aes(log(gdpPercap), lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.5, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  facet_wrap(~continent, scales = "free") +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear')
```

### But I want to do more!
Layout stretch goal: get table and figure side-by-side. This gist might get you started.

```{r}
library(ggplot2)
suppressMessages(library(dplyr))
#devtools::install_github("jennybc/gapminder")
library("gapminder")
gtbl <- tbl_df(gapminder)
```

```{r}
tbl1 <- gtbl %>%
  group_by(continent) %>% 
  summarise_each(funs(mean, median), gdpPercap) 
```
<div class="twoC">
```{r results = 'asis'}
knitr::kable(tbl1)
```

```{r}
ggplot(gtbl, aes(x = continent, y = gdpPercap, color = continent)) +
  geom_point() + ggtitle("GDP per cap by continent") 
```
</div>
<div class="clearer"></div>

It's not working perfect for md file. However, it works the html file you can refer the html file (attached in the directory) for the correct disply. 

